%% This is file `jcomp-template.tex',
%% 
%% Copyright 2017 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%%
%% $Id: jcomp-template.tex 100 2017-07-14 13:15:12Z rishi $
%%
%% Use the option review to obtain double line spacing
%\documentclass[times,review,preprint,authoryear]{elsarticle}

%% Use the options `twocolumn,final' to obtain the final layout
%% Use longtitle option to break abstract to multiple pages if overfull.
%% For Review pdf (With double line spacing)
%\documentclass[times,twocolumn,review]{elsarticle}
%% For abstracts longer than one page.
%\documentclass[times,twocolumn,review,longtitle]{elsarticle}
%% For Review pdf without preprint line
%\documentclass[times,twocolumn,review,nopreprintline]{elsarticle}
%% Final pdf
\documentclass[times,final]{elsarticle}
%%
%\documentclass[times,twocolumn,final,longtitle]{elsarticle}
%%


%% Stylefile to load JCOMP template
\usepackage{jcomp}
\usepackage{framed,multirow}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{latexsym}

% Following three lines are needed for this document.
% If you are not loading colors or url, then these are
% not required.
\usepackage{url}
\usepackage{xcolor}
\definecolor{newcolor}{rgb}{.8,.349,.1}

\journal{Journal of Computational Physics}

\begin{document}

\verso{Given-name Surname \textit{etal}}

\begin{frontmatter}

\title{Choosing optimal partitions for massively parallel transport sweeps on unstructured grids}%\tnoteref{tnote1}}%
%\tnotetext[tnote1]{This is an example for title footnote coding.}

\author[1]{Tarek Habib \snm{Ghaddar}}%\corref{cor1}}
%\cortext[cor1]{Tarek Gahddar: 
%  Tel.: +0-000-000-0000;  
%  fax: +0-000-000-0000;}
\author[1]{Jean C. \snm{Ragusa}}
%\fntext[fn1]{This is author footnote for second author.}  

\address[1]{Department of Nuclear Engineering, Texas A\&M University, College Station, TX 77843, USA}

\received{1 May 2013}
\finalform{10 May 2013}
\accepted{13 May 2013}
\availableonline{15 May 2013}
\communicated{S. Sarkar}


\begin{abstract}
%%%
A concise and factual abstract is required. The abstract should state briefly the purpose of the research, the principal results and major conclusions. An abstract is often presented separately from the article, so it must be able to stand alone. For this reason, References should be avoided, but if essential, then cite the author(s) and year(s). Also, non-standard or uncommon abbreviations should be avoided, but if essential they must be defined at their first mention in the abstract itself.
%%%%
\end{abstract}

%\begin{keyword}
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
%\MSC 41A05\sep 41A10\sep 65D05\sep 65D17
%% Keywords
%\KWD Keyword1\sep Keyword2\sep Keyword3
%\end{keyword}

\end{frontmatter}

%\linenumbers

%% main text
\section{Introduction}

The field of radiation transport studies the distribution of radiation (particle/energy) throughout a six-dimensional phase-space consisting of time, space, energy, and direction. 
Radiation transport is commonly used in reactor physics, medical physics, radiation shielding, criticality safety applications, stellar atmospheres, as well as the general simulation of radiation in different environments. 
Radiation transport is described by the Boltzmann transport equation \cite{bell_glasstone,zweifel,davison,duderstadt} and can be solved stochastically or deterministically.

The work presented in this paper deals with the transport equation discretized using multigroup in energy, discrete-ordinates (collocation) in angle, and discontinuous finite elements in space. 
The resulting discrete system of equations is amenable to iterative techniques whereby the streaming+total interaction operator need not to be assembled, generating large memory savings while being computationally very efficient.  
Essentially, this allows for a matrix-free approach, termed ``transport sweeps'' in the radiation transport community, where small local system of equations are solved cell-by-cell, for each angular direction and energy group, in a fashion that follows the flow of radiation in the computational mesh.

It is often desirable to split the problem domain so that the memory cost can be distributed across multiple processors.
Extensive research has been done in parallelizing the transport sweep \cite{KBA,partisn,denovo,mpadams2013,mpadams2015}, as the transport sweep gains complexity for distributed meshes. Maintaining sweep performance for massively parallel simulations has and continues to be a thriving research topic.

Research on sweeping on unstructured meshes has come to prominence in recent years, as the need to simulate more complex problems has risen.
Partitioning unstructured meshes in a balanced (or near equivalent amount of work per processor) fashion that preserves mesh integrity can present challenges.

The following sections will discuss a brief history of the transport sweep, parallel transport sweeps, load balancing unstructured meshes for transport, a time-to-solution estimator, and optimizing partitions for parallel transport.

Section \ref{cha:transport_sweeps} introduces the steady-state neutron transport equation and the discretization process to obtain the discrete form of the transport equation. Section \ref{cha:transport_sweeps} also provides more details on the transport sweep.

Massively parallel transport sweeps have been shown to scale up to 1.5 million cores on logically Cartesian grids using Texas A\&M's flagship deterministic transport code, PDT \cite{mpadams2013,mpadams2015}.
Section \ref{cha:parallel_transport} describes the basics of a parallel sweep algorithm, and specifically describes the KBA algorithm \cite{KBA} and PDT's extension of the KBA algorithm.
This Section also describes PDT's performance model, which predicts the sweep time given a set of partitioning parameters.

In order to solve a wider set of problems, an unstructured meshing capability was implemented in PDT.
However, unstructured meshes may lead to imbalanced partitions, or different numbers of mesh cells per processor subdomain.
To address this, two load-balancing algorithms have been implemented in PDT, the original load-balancing algorithm and the load-balancing-by-dimension algorithm \cite{mastersthesis,mc2017}.

PDT's existing performance model does not account for unstructured meshes and imbalanced partitions.
Section \ref{cha:tts} describes the time-to-solution estimator, a graph-based method that estimates the time it takes to sweep across a problem given a mesh, partitioning scheme, and machine-specific parameters.

The time-to-solution estimator described in Section \ref{cha:tts} is used to optimize the partitioning scheme.
Section \ref{cha:optimization} details two optimization methods: a ``black box'' method that uses Python's scipy.optimize library and a ``human-intelligence'' method that prioritizes partition placement in locations that do not add cells to a mesh.

\section{The transport sweep}\label{cha:transport_sweeps}
hahahaha

\section{The parallel transport sweep}\label{cha:parallel_transport}
hahahaha

\section{Unstructured meshing and load balancing in PDT}\label{cha:lb}
hahahaha

\section{The time-to-solution estimator}\label{cha:tts}

\section{Choosing optimal partitions}\label{cha:optimization}

\section*{Acknowledgments}
Acknowledgments should be inserted at the end of the paper, before the
references, not as a footnote to the title. Use the unnumbered
Acknowledgements Head style for the Acknowledgments heading.

\section*{References}

Please ensure that every reference cited in the text is also present in
the reference list (and vice versa).

\section*{\itshape Reference style}

Text: All citations in the text should refer to:
\begin{enumerate}
\item Single author: the author's name (without initials, unless there
is ambiguity) and the year of publication;
\item Two authors: both authors' names and the year of publication;
\item Three or more authors: first author's name followed by `et al.'
and the year of publication.
\end{enumerate}
Citations may be made directly (or parenthetically). Groups of
references should be listed first alphabetically, then chronologically.

%%Vancouver style references.
\bibliographystyle{model1-num-names}
\bibliography{refs}

\end{document}

%%
